---
title: "p8105_hw3_baw2160"
author: "Bryce Wong"
date: "October 8, 2018"
output: github_document
---

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(ggridges)
library(patchwork)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```
## Problem One:

### install.packages("devtools")
```{r}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
```
### Read and clean data:
* Formatted the data to use appropriate variable names.
* Focused on the “Overall Health” topic.
* Include only responses from "Excellent" to "Poor"
* Organized responses as a factor taking levels from “Excellent” to “Poor”

```{r}
brfss_data = brfss_smart2010 %>% 
  filter(Topic == "Overall Health") %>% 
  rename(Location_State = Locationabbr, Location_County = Locationdesc, Proportion = Data_value) %>%
  mutate(
    Response = as.factor(Response)
  )
```

### Answering some questions:

* In 2002, which states were observed at 7 locations?
```{r}
states_with_7_obs = brfss_data %>%  
  filter(Year == "2002") %>%
  distinct(Location_County, .keep_all = TRUE) %>% 
  count(Location_State) %>% 
  filter(n == 7)
```

**Answer:** `r states_with_7_obs$Location_State`

* Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.
```{r}
counts_per_state = distinct(brfss_data, Location_County, .keep_all = TRUE) %>% 
  count(Location_State, Year) 
ggplot(counts_per_state, aes(x = Year, y = n, group = Location_State)) +
  geom_line(aes(color = Location_State), alpha = .5)
```

* Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r}
NY_data = brfss_data %>% 
  group_by(Location_State, Year) %>% 
  spread(key = Response, value = Proportion) %>% 
  filter(Location_State == "NY", Year %in% c("2002", "2006", "2010")) %>% 
  summarize(mean_excellent = mean(Excellent, na.rm = TRUE),
            sd_excellent = sd(Excellent, na.rm = TRUE)) %>% 
  knitr::kable(digits = 1)
NY_data
```

* For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r}
state_avgs = brfss_data %>%
  spread(key = Response, value = Proportion) %>% 
  group_by(Year, Location_State) %>% 
  janitor::clean_names() %>% 
  summarize(
    avg_excellent = mean(excellent, na.rm = TRUE), 
    avg_good = mean(good, na.rm = TRUE), 
    avg_verygood = mean(very_good, na.rm = TRUE),
    avg_fair = mean(fair, na.rm = TRUE),
    avg_poor = mean(poor, na.rm = TRUE)
  )  
gather(state_avgs, key = health_score, value = averages, avg_excellent:avg_poor) %>%
  ggplot(aes(x = Year, fill = Location_State)) +
  geom_density(alpha = .5) +
  facet_grid(~health_score) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```

##Problem 2:
The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations. 

* Size of the data: `r dim(instacart)[1]` observations, with `r dim(instacart)[2]` variables.
* Structure of the data: `r class(instacart)`. Each row of the table appears to correspond to a single purchase.
* Key variables:
  *

Then, do or answer the following (commenting on the results of each):

How many aisles are there, and which aisles are the most items ordered from?
Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

